---
title: "Lab 3"
author: "Amelia Ritger"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
```

## Binary logistic regression: Penguins!

#### Read in the data
```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

penguins <- full_join(chinstrap, gentoo) %>% #will full join work with different # variables?
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% #make entries in "sex" column lower case
  filter(sex %in% c("male", "female")) #because some entries were not entered correctly in "sex" 
  #filter(sex == "male" | sex == "female") equivalent to %in% operator
```

#### Now let's graph it
```{r}
ggplot(data=penguins, aes(x=sex, y=body_mass_g)) +
  geom_jitter(aes(color = sex, pch = sex)) +
  facet_wrap(~species)
```

#### Code Chinstrap penguins as (1) and Gentoo penguins as (0)
```{r}
unique(penguins$species)

penguin_bin <- penguins %>% 
  mutate(sp_binary = case_when(species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1, species == "Gentoo penguin (Pygoscelis papua)" ~ 0))
```

#### Now let's run our binary logistic regression
```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g, data=penguin_bin, family = binomial)
summary(penguin_blr)
```
body_mass_g estimate = -0.016 means: as body mass increases, it is becoming less probable that a penguin is Chinstrap

#### Let's make some predictions for penguins (classification...kinda):
1. What is the probability that a penguin is a chinstrap penguin, if it weighs 4500 g and is male? (hint: it's probably chinstrap)
```{r}
df_m4500 <- data.frame(sex = "male", body_mass_g = 4500) #single = means assigning a value, double == is asking a question, does this match?

# Find log odds of this penguin being Chinstrap:
m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link") #type = link is default but obvs log regression link is log odds
m4500_logodds #1:4 odds

m4500_prob <- predict(penguin_blr, newdata = df_m4500, type = "response")
m4500_prob 
```

2. What is the probability that a penguin is Chinstrap, if it is recorded as a female of 4000 g?
```{r}
df_f4000 <- data.frame(sex = "female", body_mass_g = 4000) #single = means assigning a value, double == is asking a question, does this match?

# Find log odds of this penguin being Chinstrap:
f4000_logodds <- predict(penguin_blr, newdata = df_f4000, type = "link") #type = link is default but obvs log regression link is log odds
f4000_logodds #1:4 odds

f4000_prob <- predict(penguin_blr, newdata = df_f4000, type = "response")
f4000_prob
```

#### Now let's make an entire data frame and use it to make then visualize logistic regression outcomes
```{r}
penguins_mock <- data.frame(body_mass_g = rep(seq(3000, 6000, length = 200),2),#200 equally spaced observations with a range 3000-6000, done twice because use one set for female, one set for male
                            sex = c(rep("male", 200), rep("female", 200))) #"male" listed 200 times, then "female" 200 times
```

#### Now let's use those to make predictions for all of them:
```{r}
full_predict <- predict(penguin_blr, newdata = penguins_mock, type = "response", se.fit = TRUE)

#Get those into a data frame, along with the mock data used to create them

final_df <- data.frame(penguins_mock, full_predict$fit, full_predict$se.fit)
colnames(final_df) <- c("penguin_mass", "sex", "probability", "se")
```

#### Now let's graph it!
```{r}
ggplot(data = final_df, aes(x=penguin_mass, y=probability)) +
  geom_line(aes(color = sex)) +
  geom_ribbon(aes(ymin = probability-se, ymax = probability+se, fill=sex), alpha=0.3) #show standard error
```


## NOW some spatial data with sf

#### Read in the spatial data with read_sf() and practice filtering
```{r}
cougars <- read_sf(dsn = here("cougar_connections"), layer = "ds1014") %>%
  clean_names()

large_corridors <- cougars %>%
  filter(area_ac > 20000) %>%  #look at rows where area_ac>20000
  select(area_ac) #only keep that column

plot(large_corridors)
```

```{r}
cougar_sub <- cougars %>% 
  select(elev_mean)
```

```{r}
ggplot(data = cougar_sub) +
  geom_sf(aes(fill = elev_mean), color = NA)
```

#### make it interactive!
```{r}
tmap_mode("view") #view is interactive

tm_shape(cougar_sub) +
  tm_fill("elev_mean")
```


